---
title: "challengeABM: Within Herd Transmission"  
description: "A vignette simulating an FMDV introduction to a cattle herd"
format: 
   gfm:
     toc: true
     toc-depth: 2
     code-fold: show
     code-summary: "Hide code"
     preview-mode: raw
editor_options: 
  chunk_output_type: console
---

## Description  
The **challangeABM** package includes the *simulate_within_herd()* function that is calibrated to exhibit the within host virus dynamics as the *simulate_room_to_room()* function uses for laboratory based animal experiments.  At simulation start, cattle (agents) are populated to a common area with a specified number of FMDV infected individuals.  On simulation hour 1, cattle begin making random contact with each other at a specified hourly rate to show homogeneous mixing.  Depending on the quantity of virus in the donor cattle at the time of contact, they may or may not infect others.  The quantity of virus in nasal passages and in blood serum has been calibrated to replicate like Foot-and-Mouth Disease Virus (FMDV).  
 
The status of individual cattle virus loads, clinical presentation, and other factors are recorded on an hourly basis. 
     
Virus and cattle parameters may be specified in configuration files to simulate changes to infectivity and virulence.  
   
```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(here)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(pals)
```

## Load challengeABM
```{r}
library(challengeABM)
```

## Run the simulation 
First, set random seed for reproducible runs.   
```{r}
seed_val <- round(runif(1, 100, 300),0)
seed_val 
set.seed(seed_val)
```
  
Run the simulation based using a configuration file with needed parameters..      
```{r}
model_result <- simulate_within_herd(here("config/base_config_fmdv.yaml"))
```

## Within-Host Dynamics  
Randomly select 21 animals.
```{r fig.width=10, fig.height=10}
plot_individual_virus_loads(model_result, sample_n = 21)
```

## Herd Outbreak Curve     
Compare the number of infected and recovered individuals.  
```{r fig.width=8, fig.height=8}
plot_herd_curve(model_result)
```
